True versus apparent shapes of bow shocks

### 2018-01-20 papers/Papers1234/app-intermediate.tex
Transformation of basis vectors, but not sure this is right approach
Need to transform the points or the basis, but not both

The transformation from the body frame (unprimed) to observer-frame
(primed) coordinates is a rotation about the \(y\) axis by an angle
\(i\), which is described by the rotation matrix:
\begin{equation}
  \label{eq:Rotation-matrix-y}
  \mathbfss{A}_y(i) = 
   \begin{pmatrix}
    \cos i & 0 & -\sin i \\
    0 & 1 & 0 \\
    \sin i & 0 &\cos i 
  \end{pmatrix} \ .
\end{equation}
This is used in equation~\eqref{eq:Trans}.

Since we are considering cylindrically symmetric bows, all azimuths
\(\phi\) are equivalent, so it is sufficient to work with two-dimensional
curves in the plane \(z = 0\) (which is also \(\phi = 0\)) and then find
the three-dimensional surface by rotating about the \(x\)-axis via the
rotation matrix:
\begin{equation}
  \label{eq:Rotation-matrix-x}
  \mathbfss{A}_x(\phi) = 
   \begin{pmatrix}
    1 & 0 & 0 \\
    0 &\cos\phi & -\sin\phi \\
    0 &\sin\phi & \cos\phi 
  \end{pmatrix} \ .
\end{equation}
This is used in \S~\ref{sec:unit-vectors-normal} to find the general
form of the body-frame normal and tangent unit vectors:
\begin{align}
  \uvec{n} &= \mathbfss{A}_x(\phi) \, \uvec{n}_0 =
  \begin{Vector}
    \sin\alpha \\ \cos\alpha\cos\phi \\ \cos\alpha\sin\phi
  \end{Vector} \\
  \uvec{t} &= \mathbfss{A}_x(\phi) \, \uvec{t}_0 =
  \begin{Vector}
    -\cos\alpha \\ \sin\alpha\cos\phi \\ \sin\alpha\sin\phi
  \end{Vector} \ . 
\end{align}

>>>>

The transformation from the body frame (unprimed) to observer-frame
(primed) coordinates is a rotation about the \(y\) axis by an angle
\(i\), which is described by the rotation matrix:
\begin{equation}
  \label{eq:Rotation-matrix-y}
  \mathbfss{A}_y(i) = 
   \begin{pmatrix}
    \cos i & 0 & -\sin i \\
    0 & 1 & 0 \\
    \sin i & 0 &\cos i 
  \end{pmatrix} \ .
\end{equation}
This is used in equation~\eqref{eq:Trans}.  A further application is
to find the observer-frame Cartesian basis vectors by rotating the
body-frame basis:
\begin{align}
  \label{eq:xunit-prime}
  \uvec{x}' &= \mathbfss{A}_y(i)\, \uvec{x}
              =
              \begin{Vector}
              \cos i \\ 0 \\ \sin i  
              \end{Vector}
  \\
  \label{eq:yunit-prime}
  \uvec{y}' &= \mathbfss{A}_y(i)\, \uvec{y}
              =
              \begin{Vector}
              0 \\ 1 \\ 0
              \end{Vector}
  \\
  \label{eq:zunit-prime}
  \uvec{z}' &= \mathbfss{A}_y(i)\, \uvec{z}
              =
              \begin{Vector}
              -\sin i \\ 0 \\ \cos i  
              \end{Vector}
\end{align}

Since we are considering cylindrically symmetric bows, all azimuths
\(\phi\) are equivalent, so it is sufficient to work with two-dimensional
curves in the plane \(z = 0\) (which is also \(\phi = 0\)) and then find
the three-dimensional surface by rotating about the \(x\)-axis via the
rotation matrix:
\begin{equation}
  \label{eq:Rotation-matrix-x}
  \mathbfss{A}_x(\phi) = 
   \begin{pmatrix}
    1 & 0 & 0 \\
    0 &\cos\phi & -\sin\phi \\
    0 &\sin\phi & \cos\phi 
  \end{pmatrix} \ .
\end{equation}
This is used in \S~\ref{sec:unit-vectors-normal} to find the general
form of the body-frame normal and tangent unit vectors:
\begin{align}
  \uvec{n} &= \mathbfss{A}_x(\phi) \, \uvec{n}_0 =
  \begin{Vector}
    \sin\alpha \\ \cos\alpha\cos\phi \\ \cos\alpha\sin\phi
  \end{Vector} \\
  \uvec{t} &= \mathbfss{A}_x(\phi) \, \uvec{t}_0 =
  \begin{Vector}
    -\cos\alpha \\ \sin\alpha\cos\phi \\ \sin\alpha\sin\phi
  \end{Vector} \ . 
\end{align}
These are then to the observer frame in equations~\eqref{eq:nprime}
and \eqref{eq:tprime} by use of equation~\eqref{eq:Rotation-matrix-y}.

The first equality in equation~\eqref{eq:tanphi} is found by
substituting equations~\eqref{eq:nprime} and~\eqref{eq:zunit-prime}
in~\eqref{eq:tangent-line-condition}, yielding

### 2018-01-20 papers/Papers1234/sec-projection.tex
Transformation of basis vectors, but not sure this is right approach
Need to transform the points or the basis, but not both

\subsection{Unit vectors normal and tangential to the surface}
\label{sec:unit-vectors-normal}
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/bowshock-unit-vectors}
  \caption{Unit vectors in the body frame that are normal and
    tangential to the surface \(R(\theta)\) in a plane of constant
    azimuth, \(\phi\).}
  \label{fig:unitvec}
\end{figure}
We define unit vectors \(\uvec{n}, \uvec{t}\), such that \(\uvec{n}\) is normal to the surface, while \(\uvec{t}\) is tangent to the surface in a plane of constant \(\phi\). For \(\phi = 0\) the surface lies in the \(xy\) plane and it is straightforward to show (Fig.~\ref{fig:unitvec}) that in this case the unit vectors are given by 
\begin{equation}
  \label{eq:uvec-phi0}
  \uvec{t}_0 =
  \begin{Vector}
    -\cos\alpha \\ \sin\alpha \\ 0
  \end{Vector}
  \quad \mathrm{and} \quad
  \uvec{n}_0 =
  \begin{Vector}
    \sin\alpha \\ \cos\alpha \\ 0
  \end{Vector}, 
\end{equation}
where \(\alpha\) is the \textit{slope angle}, given by
\begin{equation}
  \label{eq:alpha}
  \tan\alpha = -\left.\frac{dy}{dx}\right\vert_{R(\theta)} 
  = \frac{1 + \omega \tan\theta}{\tan\theta - \omega}
\end{equation}
and \(\omega\) is a dimensionless \textit{local growth factor}:
\begin{equation}
  \label{eq:omega}
  \omega(\theta) = \frac{1}{R} \frac{dR}{d\theta} . 
\end{equation}
For general \(\phi\), we find \(\uvec{n}\) and \(\uvec{t}\) by rotating
equations~(\ref{eq:uvec-phi0}) around the \(x\)-axis with the matrix
\(\mathbfss{A}_x(\phi)\)(see Appendix~\ref{sec:plane-sky-projection}) and
then converting to the observer frame by use of
equation~(\ref{eq:Trans}) to give
\begin{gather}
  \label{eq:nprime}
  \begin{aligned}
    \uvec{n}' &= \mathbfss{A}_y(i)\, \uvec{n}
    = \mathbfss{A}_y(i)\, \mathbfss{A}_x(\phi)\, \uvec{n}_0 \\
    & =
    \begin{aligned}[t]
      & \frac{1}{(1+\omega^2)^{1/2}} \\
      & \times 
      \begin{Vector}
        (\cos\theta+\omega\sin\theta)\cos i 
        - (\sin\theta-\omega\cos\theta) \sin i \sin\phi\\
        (\sin\theta-\omega\cos\theta)\cos\phi \\
        (\cos\theta+\omega\sin\theta)\sin i 
        + (\sin\theta-\omega\cos\theta)\sin\phi\cos i
      \end{Vector}
    \end{aligned}
  \end{aligned}\\
  \label{eq:tprime}
  \begin{aligned}
    \uvec{t}' &= \mathbfss{A}_y(i)\, \uvec{t}
    = \mathbfss{A}_y(i)\, \mathbfss{A}_x(\phi)\, \uvec{t}_0 \\
    & =
    \begin{aligned}[t]
      & \frac{1}{(1+\omega^2)^{1/2}} \\
      & \times 
      \begin{Vector}
        -(\sin\theta-\omega\cos\theta)\cos i
        - (\cos\theta+\omega\sin\theta)\sin\phi\sin i \\
        (\cos\theta+\omega\sin\theta)\cos\phi \\
        -(\cos\theta+\omega\sin\theta)\sin i 
        + (\sin\theta-\omega\cos\theta)\sin\phi\cos i
      \end{Vector} \ .
    \end{aligned}
  \end{aligned}
\end{gather}
\subsection{Tangent line}
\label{sec:tangent-line}

@@

Projective space \(\mathbb{RP}^2\) or \(\mathbbm{RP}^2\) or \(\mathds{RP}^2\)

>>>>

\subsection{Unit vectors normal and tangential to the surface}
\label{sec:unit-vectors-normal}
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/bowshock-unit-vectors}
  \caption{Unit vectors in the body frame that are normal and
    tangential to the surface \(R(\theta)\) in a plane of constant
    azimuth, \(\phi\).}
  \label{fig:unitvec}
\end{figure}
We define unit vectors \(\uvec{n}, \uvec{t}\), such that \(\uvec{n}\) is normal to the surface, while \(\uvec{t}\) is tangent to the surface in a plane of constant \(\phi\). For \(\phi = 0\) the surface lies in the \(xy\) plane and it is straightforward to show (Fig.~\ref{fig:unitvec}) that in this case the unit vectors are given by 
\begin{equation}
  \label{eq:uvec-phi0}
  \uvec{t}_0 =
  \begin{Vector}
    -\cos\alpha \\ \sin\alpha \\ 0
  \end{Vector}
  \quad \mathrm{and} \quad
  \uvec{n}_0 =
  \begin{Vector}
    \sin\alpha \\ \cos\alpha \\ 0
  \end{Vector}, 
\end{equation}
where \(\alpha\) is the \textit{slope angle}, given by
\begin{equation}
  \label{eq:alpha}
  \tan\alpha = -\left.\frac{dy}{dx}\right\vert_{R(\theta)} 
  = \frac{1 + \omega \tan\theta}{\tan\theta - \omega}
\end{equation}
and \(\omega\) is a dimensionless \textit{local growth factor}:
\begin{equation}
  \label{eq:omega}
  \omega(\theta) = \frac{1}{R} \frac{dR}{d\theta} . 
\end{equation}
For general \(\phi\), we find \(\uvec{n}\) and \(\uvec{t}\) by rotating
equations~(\ref{eq:uvec-phi0}) around the \(x\)-axis with the matrix
\(\mathbfss{A}_x(\phi)\) and then rotating to the observer frame with
\(\mathbfss{A}_y(i)\)

by use of
equation~(\ref{eq:Trans}) to give
\begin{equation}
  \label{eq:nprime}
  \begin{aligned}
    \uvec{n}' &= \mathbfss{A}_y(i)\, \uvec{n}
    = \mathbfss{A}_y(i)\, \mathbfss{A}_x(\phi)\, \uvec{n}_0 \\
    & =
    \begin{aligned}[t]
      & \frac{1}{(1+\omega^2)^{1/2}} \\
      & \times 
      \begin{Vector}
        (\cos\theta+\omega\sin\theta)\cos i 
        - (\sin\theta-\omega\cos\theta) \sin i \sin\phi\\
        (\sin\theta-\omega\cos\theta)\cos\phi \\
        (\cos\theta+\omega\sin\theta)\sin i 
        + (\sin\theta-\omega\cos\theta)\sin\phi\cos i
      \end{Vector}
    \end{aligned}
  \end{aligned}
\end{equation}
\begin{equation}
  \label{eq:tprime}
  \begin{aligned}
    \uvec{t}' &= \mathbfss{A}_y(i)\, \uvec{t}
    = \mathbfss{A}_y(i)\, \mathbfss{A}_x(\phi)\, \uvec{t}_0 \\
    & =
    \begin{aligned}[t]
      & \frac{1}{(1+\omega^2)^{1/2}} \\
      & \times 
      \begin{Vector}
        -(\sin\theta-\omega\cos\theta)\cos i
        - (\cos\theta+\omega\sin\theta)\sin\phi\sin i \\
        (\cos\theta+\omega\sin\theta)\cos\phi \\
        -(\cos\theta+\omega\sin\theta)\sin i 
        + (\sin\theta-\omega\cos\theta)\sin\phi\cos i
      \end{Vector} \ .
    \end{aligned}
  \end{aligned}
\end{equation}

\subsection{Tangent line}
\label{sec:tangent-line}

@@
